\section{Academic Background}
Over my years at Miami University, I explored a variety of majors and fields. My first major, declared as a freshman, was mathematics. I enjoyed high-level mathematics and statistics courses, while also exploring a variety of subjects such as economics and organic chemistry. I also took numerous courses in psychology, which unintentionally led to my completion of a second major in psychology. Throughout my exploration of these diverse subjects, I have been most challenged and stimulated by those that required strong analytic thinking and problem solving. For that reason, I continued to take courses in mathematics and statistics even after I had completed the requirements for my major.I had not had any experience with programming prior to taking a mandatory computer science course at the end of my senior year, in the spring of 2013. I took an introductory class for beginning computer science and engineering students. Taking this class changed the course my academic life. Although I had enjoyed many courses, I had never before felt such a passion for what I was learning. My professor suggested that I take a few more courses in the summer to further evaluate my interest and aptitude, and then to consider applying to graduate school. I took his advice, deferred graduation and took two more computer science courses, Object-Oriented Programing and Data Structures, over the summer.  My interest in the subject had only grown. My intention at the end of the summer was to get a minor in computer science, and apply to graduate school in computer science. I spoke with the Graduate Director about  which computer science courses I should be taking to prepare myself for applying to graduate school. He recommended applying to the combined program, as that would make more sense for someone in my particular circumstances. So, I spent the 2013-2014 academic year taking graduate courses in computer science. These courses were challenging, especially those that were more coding intensive. However, I found that I really enjoyed the theory-based courses such as Algorithms and Automata, as they were strongly linked to my background in mathematics. I think that my experience in the Masters of Computer Science program would equate to an accelerated undergraduate program, as I learned a lot of background information in a very short period of time. Additionally, my work for my thesis has given me a lot of experience in computer science research.